---
title: "Dashboard Workshop"
output: 
  flexdashboard::flex_dashboard:
    runtime: shiny
    orientation: rows
---

```{r setup, include=FALSE}
#^^^NOTE the different header above^^^
#This is because I use a reactive environment on page 3 now

#Incorporating Shiny (interactive) components into flexdashboard:
#https://rmarkdown.rstudio.com/flexdashboard/shiny.html#simple_example
library(flexdashboard)
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(reshape2)
```
Data Overview
===================================== 

Row {data-height=150}
-----------------------------------------------------------------------

### Introduction

In this dashboard we will explore a very dull and over-used data set called iris. It's okay that it's boring; it allows us to focus on learning a new tool, which is flexdashboard! Below you see some eye-catching plots. These, on the first page, will typically be high-level summaries of the data. 


Row {data-height=400}
-----------------------------------------------------------------------

### Distribution of Petal Length by Species

```{r, fig.width = 15}
 ggplot(data = iris) +
      geom_density(aes(x = Petal.Length, fill = Species)) +
  facet_wrap(~Species, nrow = 1)+
  scale_fill_brewer(palette = "Greens") +
  theme_bw()

```


Row {data-height=400}
-----------------------------------------------------------------------

### Relationship between Sepal Width, Length

```{r}
 ggplot(data = iris) +
      geom_point(aes(x=Sepal.Width, y = Sepal.Length, colour = Species))+
      scale_colour_brewer(palette = "Greens")+
  theme_bw()

```

### Relationship between Petal Width, Length

```{r}
 ggplot(data = iris) +
      geom_point(aes(x=Petal.Width, y = Petal.Length, colour = Species))+
      scale_colour_brewer(palette = "Greens")+
  theme_bw()
```


Prediction tools: at a glance
===================================== 

Row {data-height=150}
-----------------------------------------------------------------------

### Classification tree

Our primary goal is to predict the species based on physical characteristics. We've fit a classification tree and random forest. On the left you'll see a visual of the tree's decision rules. On the right, you'll see the variance importance plot from the random forest. I made no effort to tune either of these tools. Nor did I split the data into training/testing for evaluation. It's good to be transparent about these things.

Row
-----------------------------------------------------------------------

```{r, echo=FALSE}
#setup code!
library(rpart)
library(rpart.plot)
library(randomForest)
mytree <- rpart(Species ~., data = iris)

myforest <- randomForest(Species ~., data = iris, importance = TRUE)
```

### Accuracy (in-sample)

```{r}
#https://ionicons.com/v2/cheatsheet.html
#obviously, we know in-sample measures of error are meaningless,
iris$predicted <- predict(myforest, iris, type = "class")
library(caret)
cm <- confusionMatrix(iris$predicted, iris$Species)

accuracy <- cm$overall[1]
valueBox(accuracy, 
           icon = "database",
           color = ifelse(accuracy > .8, "blue", "orange")) #conditional coloring!


```

### Setosa Sensitivity

```{r}
#obviously, we know in-sample measures of error are meaningless,



setosa_sensitivity <-cm$byClass[1,1]
valueBox(setosa_sensitivity, 
           icon = "equals",
           color = "teal")


```

### Versicolor Sensitivity

```{r}
#obviously, we know in-sample measures of error are meaningless,
vs_sensitivity <-cm$byClass[2,1]
valueBox(accuracy, 
           icon = "equals",
           color = "navy")


```


Row
-----------------------------------------------------------------------
### Classification tree

```{r}


rpart.plot(mytree)

```

### Ranking of important factors for predicting species

```{r}

varImpPlot(myforest, type = 1)

```

Make a prediction
===================================== 

Inputs {.sidebar}
-----------------------------------------------------------------------

```{r}
# shiny inputs defined here - users see these
# note there are other kinds of inputs such as drop down menus, etc.
sliderInput("Sepal.Length.input", label = "New Sepal Length:",
            min = 4.300, max = 7.900, value = 4.5, step = 0.1)
sliderInput("Sepal.Width.input", label = "New Sepal Width:",
            min = 2.000, max =4.400, value = 2.5, step = 0.1)
sliderInput("Petal.Length.input", label = "New Petal Length:",
            min = 1.000, max = 6.9, value = 3, step = 0.1)
sliderInput("Petal.Width.input", label = "New Petal Width:",
            min = 0.100, max = 2.5, value = 1.3, step = 0.1)

#these inputs get saved a reactive object where you can call them as:
#for example, input$Petal.Width.input or input$Sepal.Length.input (as defined above)

```


Row 
-----------------------------------------------------------------------

### Prediction with given inputs

```{r }

new.df <- reactive({data.frame(Sepal.Length = input$Sepal.Length.input, #<-- calling the reactive input!
                     Sepal.Width = input$Sepal.Width.input,
                     Petal.Length = input$Petal.Length.input,
                     Petal.Width = input$Petal.Width.input)})

prediction_class <- reactive({predict(myforest, new.df(), type = "class")}) #based on user-supplied input!

renderValueBox({

valueBox(prediction_class(), 
           icon = "chart-line",
           color = "blue") 
  
})

```

Row 
-----------------------------------------------------------------------
### Predicted Probabilities of Species

```{r, fig.width = 7 }
  new.df_reac <-reactive({ data.frame(Sepal.Length = input$Sepal.Length.input,
                     Sepal.Width = input$Sepal.Width.input,
                     Petal.Length = input$Petal.Length.input,
                     Petal.Width = input$Petal.Width.input)})

prediction_probability_reac <- reactive({
  #notice I'm calling the reactive new.df_reac using "new.df_reac()"!
  predict(mytree, new.df_reac(), type = "prob") %>% melt
})


renderPlot({

ggplot(data =prediction_probability_reac()) + #again! using the "()" syntax to denote reactive object
  geom_bar(aes(x = Var2, y = value, fill = Var2), stat = "identity") +
  labs( x = "Species", y = "Predicted Probability") +
  theme_bw() + 
  scale_fill_brewer("Species", palette = "Greens")
  
})

```

