---
title: "Food Dashboard"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
---
```{r setup, include=FALSE}
#multiple pages within one dashboard
#https://rmarkdown.rstudio.com/flexdashboard/layouts.html#multiple_pages

# Always start with a fresh environment
rm(list = ls())

source("Ryan_Code/data_cleaning.R")


library(flexdashboard)
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(reshape2)
library(randomForest)
library(caret)
library(e1071)
library(MASS)
library(gridExtra)
library(pscl)
library(boot)
library(rpart)
library(rpart.plot)
library(rgeos)
library(tigris)

cps_fexpend <- cps[!is.na(cps$fexpend),]

cps_fexpend = subset(cps_fexpend, select = -c(id, weight, fsecurity))

cps_fsecurity <- cps[!is.na(cps$fsecurity),]

cps_fsecurity = subset(cps_fsecurity, select = -c(id, weight, fexpend))

# ANALYSIS OF DISABILITY VARIABLE

cps_fsecurity$disability_cat = ifelse(cps_fsecurity$disability > 0, "Yes", "No")
cps_fsecurity$disability_cat = as.factor(cps_fsecurity$disability_cat)

cps_fsecurity$fsecurity_cat = ifelse(cps_fsecurity$fsecurity > 0, "yes", "no")
cps_fsecurity$fsecurity_cat = as.factor(cps_fsecurity$fsecurity_cat)
#NOTE: "-----------------------------------------------------------------------" is the section delimiter
#NOTE: "===================================== " is the PAGE delimiter

```

Data Overview{data-icon="fa-signal"}
============================

Row {data-height=200}
-----------------------------------------------------------------------

### Goal of Research:
<font size = 2>
The goal of this analysis was to determine the affects certain variables have on food insecurity and food expenditure. The goal of creating predictions for food insecurity and food expenditure on a block level for the U.S. was also a goal.
</font>

### Process:
<font size = 2>
In order to accomplish said goal we will begin by performing an exploratory analysis on all of the most important variables as determined by the variable importance plot. We will then proceed to demonstrate the effects of said variables on the response variables food security and food expenditure using regression analysis. From there we will use a random forest, which is a form of prediction via machine learning, to create predictions for both response variables for national block groups.
</font>

Row
--------------
### Key Variables
<font size = 2>
Response Variables:

* fsecurity: The number of positively-answered questions from a food insecurity survey.
* fexpend: The weekly food expenditure per person in USD.

Explanatory Variables:

* Hhsize: Number of family members within a household
* Elderly: Number of elderly family members within a household
* Hispanic: Number of hispanic family members within a household
* Black: Number of black family members within a household
* Kids: Number of kids within a household
* Educated: Number of family members with an associates degree or higher
* Employed: Number of family members who are employed
* Married: Number of family members within the household who are married
* Disabled: If the household has one or more family members who are disabled. 

</font>

Row
-----------------------
### Analysis of Response variables
```{r, fig.width = 15, fig.height = 8}

a1 <- ggplot(data = cps_fsecurity, aes (x = fsecurity))+geom_bar() + geom_text(stat = 'count',aes(label=..count..), vjust = -1) + labs(x = "Food Insecurity Level - Number of positively-answered food insecurity questions", y = "Number of Households")

a2 <- ggplot(data = cps_fexpend, aes (x = fexpend, fill = fexpend))+geom_histogram(binwidth = 5) + labs(x = "Food expense - weekly food expenditure per person", y = "Number of Households") + scale_x_continuous(labels=scales::dollar_format()) + scale_fill_brewer(palette = "BuGn")

grid.arrange(a1, a2, nrow = 1)
```




Exploratory Analysis - Food Security {data-navmenu="Exploratory Analysis"}
===============

Row
--------------------
### Description:
<font size = 3>
These plots give a physical representation of the connections between individual important variables and the response variable food security.
</font>

Column {.tabset}
---------------
### Disability Analysis 
```{r, fig.width = 20, fig.height = 10}
p1 <- ggplot(data = cps_fsecurity, aes(x = disability_cat, fill = disability_cat))+geom_bar() + geom_text(stat = 'count',aes(label=..count..), vjust = -1) + 
  labs(x = "Disabled Individual Living Within Household", y = "Number of Households", fill = "If Disabled") + scale_fill_brewer(palette = "Blues")

cps_disability <- cps_fsecurity %>% group_by(disability_cat) %>% summarise(Average = mean(fsecurity))

p2 <- ggplot(aes(x = disability_cat, y = Average, fill = Average), data = cps_disability ) + geom_bar(stat = "Identity") +
  labs(x = "Disabled Individual Living Within Household", y = "Average Level of Food Insecurity", fill = "Average Level") +
  scale_fill_distiller(palette = "Blues")

p3 <- ggplot() + geom_boxplot(aes(group = disability_cat, x = disability_cat, y = fsecurity, fill = disability_cat), data = cps_fsecurity) +
  labs(x = "Disabled Individual Living Within Household", y = "Level of Food Insecurity - log scale", fill = "If Disabled") + 
  scale_fill_brewer(palette = "Blues") + scale_y_log10() 

grid.arrange(p1, p2, p3, nrow = 1)
```

### Elderly Analysis
```{r, fig.width = 20, fig.height = 10}
g1 <- ggplot(data = cps_fsecurity, aes(x = elderly, fill = elderly))+geom_bar() + geom_text(stat = 'count',aes(label=..count..), vjust = -1) + 
  labs(x = "Number of Elderly in Household", y = "Number of Households") + scale_fill_fermenter(palette = "Blues") 

cps_elderly <- cps_fsecurity %>% group_by(elderly) %>% summarise(meld = mean(fsecurity))

g2 <- ggplot(aes(x = elderly, y = meld, fill = elderly), data = cps_elderly) + geom_bar(stat = "Identity") + 
  labs(x = "Number of Elderly in Household", y = "Average Level Of Food Insecurity", fill = "Number of Elderly") +
  scale_fill_distiller(palette = "Blues")

g3 <- ggplot() + geom_boxplot(aes(group = elderly, x = elderly, y = fsecurity, fill = elderly), data = cps_fsecurity) + 
  scale_y_log10() + scale_fill_distiller(palette = "Blues") + labs(x = "Number of Elderly Within Household", y = "Food Security Level Per Household - Log Scale", fill = "Number of Elderly")


grid.arrange(g1, g2, g3, nrow = 1)
```

### Education Analysis
```{r, fig.width = 20, fig.height = 10}
q1 <- ggplot(data = cps_fsecurity, aes(x = education))+ geom_bar() + geom_text(stat = 'count',aes(label=..count..), vjust = -1) +
  labs(x = "Number of Educated Individuals Within Household", y = "Number of Households")

cps_education <- cps_fsecurity %>% group_by(education) %>% summarise(med = mean(fsecurity))

q2 <- ggplot(aes(x = education, y = med, fill = education), data = cps_education) + geom_bar(stat = "Identity") +
  labs(x = "Number of Educated Individuals Within Household", y = "Average Level of Food Insecurity", fill = "Number of Educated") +
  scale_fill_distiller(palette = "Blues")

q3 <- ggplot() + geom_boxplot(aes(group = education, x = education, y = fsecurity, fill = education), data = cps_fsecurity) +
  scale_y_log10() + scale_fill_distiller(palette = "Blues") + labs(x = "Number of Individuals With Associates or Higher Within Household", y = "Food Security Level Per Household - log scale", fill = "Number of Educated")


grid.arrange(q1, q2, q3, nrow = 1)
```

### Employed Analysis
```{r, fig.width = 20, fig.height = 10}
p1 <- ggplot(data = cps_fsecurity, aes(x = employed))+ geom_bar() + geom_text(stat = 'count', aes(label = ..count..), vjust = -1) +
  labs(x = "Number of Employed Individuals Within Household", y = "Number of Households")

cps_employed <- cps_fsecurity %>% group_by(employed) %>% summarise(memp = mean(fsecurity))

p2 <- ggplot(aes(x = employed, y = memp, fill = employed), data = cps_employed) + geom_bar(stat = "Identity") +
  labs(x = "Number of Employed Individuals Within Household", y = "Average Level of Food Insecurity", fill = "Number of Employed") + scale_fill_distiller(palette = "Blues")
  
p3 <- ggplot() + geom_boxplot(aes(group = employed, x = employed, y = fsecurity, fill = employed), data = cps_fsecurity) +
  scale_y_log10() + scale_fill_distiller(palette = "Blues") + labs(x = "Number of Employed Individuals Within Household", y = "Food Security Level Per Household - log scale", fill = "Number of Employed")

grid.arrange(p1, p2, p3, nrow = 1)
```

### Household Size Analysis
```{r, fig.width = 20, fig.height = 10}
k1 <- ggplot(data = cps_fsecurity, aes(x = round(hhsize,0))) + geom_bar() + geom_text(stat = 'count', aes(label = ..count..), vjust = -1) +
  labs(x = "Number of Family Members Within Household", y = "Number of Households")

grid.arrange(k1, nrow = 1)
```


Exploratory Analysis - Food Expenditure {data-navmenu="Exploratory Analysis"}
==================

Row
--------------------
### Description:
<font size = 3>
These plots give a physical representation of the connections between individual important variables and the response variable food expenditure.
</font>

Column {.tabset}
----------------------
```{r, warning=FALSE, show=FALSE}
cps_fexpend_f <- cps_fexpend

cps_fexpend_f$disability <- ifelse(cps_fexpend_f$disability > 0, "Yes", "No")

breaks <- c(0, 1, 2, 3, 4, 5, 6, 7, 8 ,9 ,10, 11, 12, 13)

tags <- c('0','1', '2', '3', '4', '5', '6', '7', '8' ,'9' ,'10', '11', '12')

cps_fexpend_f$hhsize_f <- cut(cps_fexpend_f$hhsize, breaks = breaks, include.lowest = TRUE, right = FALSE, labels = tags)

```


### Disability Analysis
```{r, fig.width = 20, fig.height = 10}
f1 <- ggplot(data = cps_fexpend_f, aes (x = disability, fill = disability))+geom_bar() + geom_text(stat = 'count',aes(label=..count..), vjust = -1) + 
  labs(x = "If Disabled Individual Within Household", y = "Number of Households", fill = "If Disabled") + scale_fill_brewer(palette = "BuGn")

cps_fexpend_disability <- cps_fexpend_f %>% group_by(disability) %>% summarise(me = mean(fexpend))

f2 <- ggplot(aes(x = disability, y = me, fill = me), data = cps_fexpend_disability) + geom_bar(stat = "Identity") +
  labs(x = "Number of Disabled Individuals in Household", y = "Average Level of Food Insecurity", fill = "Average Amount in USD") +
  scale_y_continuous(labels=scales::dollar_format()) + scale_fill_distiller(palette = "BuGn")

f3 <- ggplot() + geom_boxplot(aes(group = disability, x = disability, y = fexpend, fill = disability), data = cps_fexpend_f) +
  labs(x = "If Disabled Person Lives Within Household", y = "Level of Food Expenditure Within Household", fill = "Disability Level") +
  scale_y_continuous(labels=scales::dollar_format()) + scale_fill_brewer(palette = "BuGn")

grid.arrange(f1, f2, f3, nrow = 1)
```


### Elderly Analysis
```{r, fig.width = 20, fig.height = 10}
d1 <- ggplot(data = cps_fexpend, aes (x = elderly))+geom_bar() + geom_text(stat = 'count',aes(label=..count..), vjust = -1) + 
  labs(x = "Elderly In Household", y = "Number of Households") +
  scale_fill_brewer(palette = "BuGn")

cps_fexpend_elderly <- cps_fexpend %>% group_by(elderly) %>% summarise(mexp = mean(fexpend))

d2 <- ggplot(aes(x = elderly, y = mexp, fill = mexp), data = cps_fexpend_elderly) + geom_bar(stat = "Identity") + 
  labs(x = "Number of Elderly in Household", y = "Average Amount of Food Expenditure in Household", fill = "Average Amount in USD") + 
  scale_y_continuous(labels=scales::dollar_format()) + scale_fill_distiller(palette = "BuGn")

d3 <- ggplot() + geom_boxplot(aes(group = elderly, x = elderly, y = fexpend, fill = elderly), data = cps_fexpend) +
  labs(x = "Number of Elderly in Household", y = "Amount of Food Expenditure in Household - log scale", fill = "# of Elderly") +
  scale_y_continuous(labels=scales::dollar_format()) + scale_y_log10(labels = scales::dollar_format()) + 
  scale_fill_distiller(palette = "BuGn")

grid.arrange(d1, d2, d3, nrow = 1)
```

### Education Analysis
```{r, fig.width = 20, fig.height = 10}
z1 <- ggplot(data = cps_fexpend, aes(x = education))+ geom_bar(aes(fill = education)) + geom_text(stat = 'count',aes(label=..count..), vjust = -1) +
  labs(x = "Number of Educated Individuals Within Household", y = "Number of Households") +
  scale_fill_distiller(palette = "BuGn")

cps_fexpend_education <- cps_fexpend %>% group_by(education) %>% summarise(med = mean(fexpend))

z2 <- ggplot(aes(x = education, y = med, fill = med), data = cps_fexpend_education) + geom_bar(stat = "Identity") +
  labs(x = "Number of Educated in Household", y = "Average Amount of Food Expenditure in Household", fill = "Average Food Expense USD") +
  scale_y_continuous(labels=scales::dollar_format()) +  scale_fill_distiller(palette = "BuGn", labels=scales::dollar_format())

z3 <- ggplot() + geom_boxplot(aes(group = education, x = education, y = fexpend, fill = education), data = cps_fexpend) +
  scale_y_log10(labels = scales::dollar_format()) + labs(x = "Number of Educated in Household", y = "Average Amount of Food Expenditure in Househod - log scale", fill = "Number of Educated") + scale_fill_distiller(palette = "BuGn")

grid.arrange(z1, z2, z3, nrow = 1)
```


### Employed Analysis
```{r, fig.width = 20, fig.height = 10}
j1 <- ggplot(data = cps_fexpend, aes(x = employed))+ geom_bar() + geom_text(stat = 'count', aes(label = ..count..), vjust = -1) +
  labs(x = "Number of employed Individuals Within Household", y = "Number of Households")

cps_fexpend_employed <- cps_fexpend %>% group_by(employed) %>% summarise( Average = mean(fexpend))

j2 <- ggplot(aes(x = employed, y = Average, fill = Average), data = cps_fexpend_employed) + geom_bar(stat = "Identity") +
  labs(x = "Number of Employed in Household", y = "Average Amount of Food Expenditure Per Household", fill = "Average Amount in USD") +
  scale_y_continuous(labels=scales::dollar_format()) + scale_fill_distiller(palette = "BuGn")

j3 <- ggplot() + geom_boxplot(aes(group = employed, x = employed, y = fexpend, fill = employed), data = cps_fexpend) + 
  labs(x = "Number of Employed Within Household - log scale", y = "Food Expenditure Per Household", fill = "Number of Employed") +
  scale_y_log10(labels = scales::dollar_format()) + scale_fill_distiller(palette = "BuGn")

grid.arrange(j1, j2, j3, nrow = 1)
```


### Household Size Analysis
```{r, fig.width = 20, fig.height = 10}
s1 <- ggplot(data = cps_fsecurity, aes(x = hhsize)) + geom_histogram(binwidth = 1) + 
  #geom_text(stat = 'count', aes(label = ..count..), vjust = -1) +
  labs(x = "Number of Individuals Within Household", y = "Number of Households")

cps_fexpend_employed <- cps_fexpend_f %>% group_by(hhsize_f) %>% summarise(Average = mean(fexpend))

s2 <- ggplot(aes(x = hhsize_f, y = Average, fill = Average), data = cps_fexpend_employed) + geom_bar(stat = "Identity") +
  labs(x = "Number of Individuals Within Household", y = "Average Level of Food Insecurity") +
  scale_y_continuous(labels=scales::dollar_format()) + scale_fill_distiller(palette = "BuGn")

s3 <- ggplot() + geom_boxplot(aes(group = round(hhsize,0), x = round(hhsize,0), y = fexpend, fill = round(hhsize,0)), data = cps_fexpend) +
  labs(x = "Number of Individuals Within Household", y = "Food Expenditure Per Household", fill = "Number of Individuals") +
  scale_fill_distiller(palette = "BuGn") + scale_y_continuous(labels=scales::dollar_format())


s4 <- ggplot(data = cps_fexpend, aes(x = hhsize, y = fexpend))+geom_jitter()+ labs(x = "Number of Family Members Within Household", y = "Food Expense" ) +
  scale_y_continuous(labels=scales::dollar_format())

grid.arrange(s1, s2, s3, s4, nrow = 2)
```


Predictive Model - Food Security {data-navmenu="Predictive Models"}
===

Row
--------------
### Description:
<font size = "3">
This plot was created from the final random forest for food security, and it lists the variables in their importance in food security level prediction. A random forest is a collection of classification trees used to have a more accurate statistical model.
</font>

Row
--------------
```{r, fig.width = 11, fig.height = 8}
final_forest <- readRDS("final_forest.RDS")

varImpPlot(final_forest, type = 1)
```

Predictive Model - Food Expenditure {data-navmenu="Predictive Models"}
===========================

Row
----------------------------
### Description:
<font size = "3">
This plot was created from the final random forest for food expenditure per individual for one week, and it lists the variables in their importance in food expenditure amount prediction. A random forest is a collection of classification trees used to have a more accurate statistical model.
</font>

Row
--------------
```{r, fig.width = 11, fig.height = 8}
fexpend_final_forest <- readRDS("fexpend_final_forest.RDS")

varImpPlot(fexpend_final_forest, type = 1)
```


Descriptive Model - Food Security {data-navmenu="Descriptive Models"}
=============================================
Row{data-height=200}
---------------------------------------------
### Description:
<font size="3">
This model will not be used for the final prediction, however, it will be used as a representation of what the relationships between
the response variable and explanatory variables are. A Zero Inflation Poisson Model is a regular poisson model except that it accounts for a significant amount of zeroes within the response variable. This is necessary because a majority of houses within the U.S. will have stated that they are not food insecure and therefore be a zero.</font>

Column {.tabset}
---------------------------------------------
### Zero Inflation Poisson Model
<font size="3"> 
Summary 
</font> 
```{r, attr.output='style="max-height: 400px;"'}
fsecurity.glm2 = zeroinfl(fsecurity ~ disability + education + elderly + urban_c | disability + education + elderly + urban_c, data = cps_fsecurity)

summary(fsecurity.glm2)
```

### Exponentiated Coefficients
```{r, attr.output='style="max-height: 300px;"'}
exp(coef(fsecurity.glm2))
```

### Confidence Intervals
```{r, attr.output='style="max-height: 300px;"'}
exp(confint(fsecurity.glm2))
```

### Interpretations for Count Model
<font size = "2">

Explanatory Variables that Increase Food Insecurity:
  
* Disability Interpretation (Count Model): With all other variables held constant, for vulnerable individuals, the mean level of food insecurity increases by a factor of exp(0.259602) = 1.2964 if there is a disabled person living within the house. This means that if there is a disabled person living within the house, food insecurity of the house increases from the mean by about 30%.

Explanatory Variables that Decrease Food Insecurity:
  
* Education Interpretation (Count Model): With all other variables held constant, for vulnerable individuals, the mean level of food insecurity decreases by a factor of exp(-0.110751) = .89516 for every educated person within the house. This means that for every new educated individual living within the house, food insecurity of the house decreases from the mean by about 10.55%.

* Elderly Interpretation (Count Model): With all other variables held constant, for vulnerable individuals, the mean level of food insecurity decreases by a factor of exp(-0.175185) = .8393 for every educated person within the house. This means that for every new elderly person living within the house, food insecurity of the house decreases from the mean by about 16.04%.

* Micropolitan Interpretation (Count Model): With all other variables held constant, for vulnerable individuals, the mean level of food insecurity for households in micropolitan areas is exp(-0.230339) = .79426 times the mean level of food insecurity for the baseline large metro area. 

* Non core / Possibly Rural Interpretation (Count Model): With all other variables held constant, for vulnerable individuals, the mean level of food insecurity for households in rural areas is exp(-0.007156) = .99287 times the mean level of food insecurity for the baseline large metro area. 

</font>

### Interpretations for Zero-Inflation Model
<font size = "2">
<<<<<<< HEAD

Explanatory Variables that Increase the Probability of Being Food Insecure:

* Disability Interpretation (Zero-Inflation Model): With all other variables held constant, the odds a house is food insecure increase by a factor of exp(-1.04366) = 0.3515741 if there is a disabled person living within the household. This means that if there is a disabled person living within the household, the odds of a house having a food insecurity level of 1 or greater increases by 64.84%.

Explanatory Variables that Decrease the Probability of Being Food Insecure:
=======
Disability Interpretation (Zero-Inflation Model): With all other variables held constant, the odds a house is food *secure* change by a factor of exp(-1.0453348) = 0.3515741 (so they decrease). This means that if there is a disabled person living within the household, the odds of a house having a food insecurity level of 1 or greater increases by 64.84%.

Education Interpretation (Zero-Inflation Model): With all other variables held constant, the odds that a household is food secure change by a factor of exp(0.7795625) = 2.1805181 (so they increase). This means that for every person within the house with an associates degree or higher, the odds of a house having a food insecurity level of 1 or greater decreases by 118%.

Elderly Interpretation (Zero-Inflation Model): With all other variables held constant, the odds that a house is food secure change by a factor of exp(0.6967225) = 2.0071634 (so they almost double). This means that for every elderly person within the house, the odds of a house having a food insecurity level of 1 or greater decreases by 100%.
>>>>>>> 1bd5c4d8c60a1133325cbf7adb56d3a8ca38ba84

* Education Interpretation (Zero-Inflation Model): With all other variables held constant, the odds that a household is food insecure decrease by a factor of exp(0.77369) = 2.1805181 for every educated person living within the house. This means that for every person within the house with an associates degree or higher, the odds of a house having a food insecurity level of 1 or greater decreases by 118%.

* Elderly Interpretation (Zero-Inflation Model): With all other variables held constant, the odds that a house is food insecure decrease by a factor of exp(0.69504) = 2.0071634 for every elderly person living within the house. This means that for every elderly person within the house, the odds of a house having a food insecurity level of 1 or greater decreases by 100%.

* Micropolitan Interpretation (Zero-Inflation Model): With all other variables held constant, the odds that a house is food insecure decrease by a factor of exp(.39824) = 1.489 compared to the baseline reference large metro.This means that the odds of a house being food insecure when it is in a micropolitan area is 1.5 times the odds that a large metro household is food insecure. 
  
* Non core / Possibly Rural Interpretation (Zero-Inflation Model): with all other variables held constant, the odds that a house is food insecure decrease by a factor of exp(.02018) = 1.0204 compared to the baseline reference large metro. This means that the odds of a house being food insecure when it is in a non-core or rural area is 1.0204 times the odds that a large metro household is food insecure. The difference is not as significant as the micropolitan large metro difference. 
  
</font>

Descriptive Model - Food Expenditure {data-navmenu="Descriptive Models"}
==============

Row{data-height=200}
--------------
### Description:
<font size = "3">
This model will not be used for the final prediction, however, it will be used as a representation of what the relationships between
the response variable and explanatory variables are. A Zero Adjusted Gamma Model is a regular gamma model except that it accounts for a significant amount of zeroes within the response variable. This is necessary because a majority of houses within the U.S. have said that they have spent $0 this week per person within the household. Suspected reasons for this would either be people who are getting food from food banks, families that buy food in bulk, or they could be dependants, although this is unlikely since it is an average of a block. 
</font>

Column {.tabset}
-------------
### Zero Adjusted Gamma Model - Gamma Aspect

```{r, attr.output='style="max-height: 500px;"'}
cps_fexpend$fexpend_0 <- ifelse(cps_fexpend$fexpend == 0, 1, 0)

cps_fexpend_new <- subset(cps_fexpend, fexpend != 0)

fexpend.glm2 <- glm(fexpend ~ hhsize + elderly + employed + disability + education + urban_c, data = cps_fexpend_new, family = Gamma(link = "log"))

summary(fexpend.glm2)
```

### Zero Adjusted Gamma Model - Bernoulli Aspect
```{r, attr.output='style="max-height: 400px;"'}

fexpend.glm3 <- glm(fexpend_0 ~ hhsize + elderly + employed + disability + education + urban_c, data = cps_fexpend, family = binomial(link = "logit"))

summary(fexpend.glm3)
```

### Exponential Coefficients
``` {r, attr.output='sytle="max-height: 200px;"'}
exp(coef(fexpend.glm2))
exp(coef(fexpend.glm3))
```

### Confidence Intervals 
```{r, attr.output='sytle="max-height: 200px;"'}
exp(confint(fexpend.glm2))
exp(confint(fexpend.glm3))
```


### Interpretations for Gamma Model
<font size = "2">
  Explanatory Variables that Increase Food Expenditure:
  
*  Employed Interpretation (Gamma Model): With all other variables held constant, for individuals who purchase food, the mean amount of food expenditure increases by a factor of exp(.06672) = 1.0690010 for every additional employed person. This means that for every new employed individual within the household, the amount of money in U.S. dollars used to buy food for each individual increase by about 6.9% per person compared to the mean.

* Education Interpretation (Gamma Model): With all other variables held constant, for individuals who purchase food, the mean amount of food expenditure increases by a factor of exp(0.12464) = 1.1327362 for every additional educated person. This means that for every new educated individual within the household, the amount of money in U.S. dollars used to buy food for each individual increases by about 13.27% per person compared to the mean. 

Explanatory Variables that Decrease Food Expenditure:

*  Hhsize Interpretation (Gamma Model): With all other variables held constant, for individuals who purchase food, the mean amount of food expenditure changes by a factor of exp(-.22794) = .7961747 for every new person. This means that for every new person that lives within the house the amount of money used to  buy food for each individual decreases by about 20.4% per person compared to the mean.

* Elderly Interpretation (Gamma Model): With all other variables held constant, for individuals who purchase food, the mean amount of food expenditure changes by a factor of exp(-.04794) = .9531918 for every new elderly person. This means that for every new elderly person within the household, the amount of money in U.S. dollars used to buy food for each individual decreases by about 4.68% per person compared to the mean.

* Disability Interpretation (Gamma Model): With all other variables held constant, for individuals who purchase food, the mean amount of food expenditure changes by a factor of exp(-0.06584) = .9362819 for every new disabled person. This means that for every new disabled individual within the household, the amount of money in U.S. dollars used to buy food for each individual decreases by about 6.37% per person compared to the mean.

* Micropolitan Interpretation (Gamma Model): With all other variables held constant, for individuals who purchase food, the mean amount of food expenditure for households in micropolitan areas is exp(-0.201771) = .81728 times the mean amount of food expenditure for households in large metro areas.

* Non-core/possibly rural Interpretation (Gamma Model): With all other variables held constant, for individuals who purchase food, the mean amount of food expenditure for households in rural  areas is exp(-0.126072) = .8816 times the mean amount of food expenditure for households in large metro areas.

</font>
  

### Interpretations for Bernoulli Model
<font size = "2">
Explanatory Variables that Increase Probability of $0:
  
* Disability Interpretation (Bernoulli Model): With all other variables held constant, the odds of spending $0 changes by a factor of exp(0.25147) = 1.2859122. This means that for every new disabled person within the household, the odds of said household spending $0 on food per person increases by about 28.59%.

Explanatory Variables that Decrease Probability of $0:
  
* Hhsize Interpretation (Bernoulli Model): With all other variables held constant, the odds of spending $0 changes by a factor of exp(-0.07890) = .9241349. This means that for every new person within the household, the odds of said household spending $0 on food per person decreases by about 7.6%.

* Elderly Interpretation (Bernoulli Model): With all other variables held constant, the odds of spending $0 changes by a factor of exp(-0.01478) = .9853277. This means that for every new elderly person within the household, the odds of said household spending $0 on food per person decreases by about 1.47%. 

* Employed Interpretation (Bernoulli Model): With all other variables held constant, the odds of spending $0 changes by a factor of exp(-0.30615)= .7362754. This means that for every new employed person within the household, the odds of said household spending $0 on food per person decreases by about 26.37%.

* Education Interpretation (Bernoulli Model): With all other variables held constant, the odds of spending $0 changes by a factor of exp(-0.25484) = .7750404. This means that for every new educated person within the household, the odds of said household spending $0 on food per person decreases by about 22.50%.

* Micropolitan Interpretation (Bernoulli Model): With all other variables held constant, the odds of spending $0 decreases by a factor of 
exp(-0.052550) = .9488 compared to the baseline large metro area. This means that the odds of spending no money on food for individuals in micropolitan areas is about exp(-.05) TIMES the odds of spending no money on food for individuals in large metro areas. That is, the odds are smaller for individuals living in micropolitan areas. 

* Non-core/possibly rural Interpretation (Bernoulli Model): With all other variables held constant the odds of spending $0 decreases by a factor of exp(-0.291537) = .7471 compared to the baseline large metro area. This means that the odds of spending no money on food for individuals in non-core or rural areas is .7471 times the odds of spending no money on food for individuals in large metro areas. That is, the odds are smaller for individuals living in non-core or rural areas.

</font>
  

Food Security{data-navmenu="Heat Map Of Predictions"}
===========================

Inputs {.sidebar}
-----------------------------------------------------------------------
```{r, echo = FALSE}
final_fexpend <- read.csv( "Ryan_Data/final_acs_fexpend.csv")

final_fsecurity <- read.csv("Ryan_Data/final_acs_fsecurity.csv")

State_stuff = data.frame(state.abb, state.name)

final_fsecurity$GEOID = as.character(paste0(final_fsecurity$GEOID, substr(final_fsecurity$location, 13, 13)))

final_fexpend$GEOID = as.character(paste0(final_fexpend$GEOID, substr(final_fexpend$location, 13, 13)))

final_combined <- left_join(final_fexpend, final_fsecurity, by = "GEOID")

final_combined$GEOID <- ifelse(nchar(final_combined$GEOID) == 12, final_combined$GEOID, paste0("0",final_combined$GEOID))

selectInput("State", label = "Select State:", choices = unique(State_stuff$state.name)[order(unique(State_stuff$state.name))], selected = NULL)

```


Row
------------------------------
```{r, echo = FALSE}

shp_file <- reactive({ block_groups(state =  State_stuff$state.abb[State_stuff$state.name == input$State]) %>%
    left_join(final_combined, by="GEOID" ) %>%
    rmapshaper::ms_simplify( keep = 0.01, keep_shapes = TRUE) %>%
    #     mutate(COUNTYNAME = convert_fips_to_names(COUNTYFP, states = rep("State", nrow(shp_join)), geo_header="COUNTY"))
    mutate(labels = paste0(
      "<strong> County: </strong> ",
      GEOID, "<br/> ",
      "<strong> Food Expenditure: </strong> $",
      fexpend_predictions, "<br/> ",
      "<strong> Food Insecurity: </strong> ",
      fsecurity_predictions, "<br/> "
    ) %>%
      lapply(htmltools::HTML))
})


#for color scale
limit <- c(min(final_combined$fexpend_predictions,na.rm=TRUE), max(final_combined$fexpend_predictions,na.rm=TRUE))
limit_bin <- c(min(final_combined$fsecurity_predictions,na.rm=TRUE), max(final_combined$fsecurity_predictions,na.rm=TRUE))
pal_bin <- colorBin(
  palette = "RdBu", domain = final_combined$fsecurity_predictions,
  reverse = TRUE,
  bins = seq(0, max(final_combined$fsecurity_predictions, na.rm = TRUE) + .01, by = .20)
)

renderLeaflet({
  leaflet(shp_file(),height=500, width=1000) %>%
    addTiles() %>%
    #setView(lng = 0, lat = 30, zoom = 2) %>%
    addPolygons(
      fillColor = ~ pal_bin(fsecurity_predictions),
      stroke=FALSE,
      color = "white",
      fillOpacity = 0.6,
      label = ~labels,
      highlight = highlightOptions(
        color = "black",
        bringToFront = TRUE
      )
    )%>%
    leaflet::addLegend(
      pal = pal_bin, values = ~fsecurity_predictions,
      opacity = 0.7, title = "Average Level of\nFood Insecurity per\nHousehold",
      position = "topright"
    )
})

```

Row
--------------------

### Predictions
```{r, echo = FALSE, attr.output='style="max-height: 200px;"'}
temp1 <- reactive({ shp_file() %>% as.data.frame%>%dplyr::select(c("GEOID","location.y","fsecurity_predictions", "female.y", "hispanic.y",
                                  "black.y", "kids.y", "elderly.y", "education.y", "employed.y",
                                  "married.y", "disability.y", "urban_c.y"))%>%
                  rename(`Predicted Average` = fsecurity_predictions, `Location` = location.y, `Females per Household` = female.y,
                         `Hispanics per Household` = hispanic.y, `Blacks per Household` = black.y, `Kids per Household` = kids.y, 
                         `Elderly per Household` = elderly.y, `Educated per Household` = education.y, `Employed per Household` = employed.y,
                         `Married per Household` = married.y, `If Disabled in Household` = disability.y, `Urban level` = urban_c.y)

})
renderDataTable({temp1()})
```

Food Expenditure{data-navmenu="Heat Map Of Predictions"}
===

Inputs {.sidebar}
-----------------------------------------------------------------------

```{r, echo = FALSE}
selectInput("State2", label = "Select State:", choices = unique(State_stuff$state.name)[order(unique(State_stuff$state.name))], selected = NULL)
```


Row
----

```{r, echo = FALSE}

shp_file2 <- reactive({ block_groups(state = State_stuff$state.abb[State_stuff$state.name == input$State2]) %>%
    left_join(final_combined, by="GEOID" ) %>%
    rmapshaper::ms_simplify( keep = 0.01, keep_shapes = TRUE) %>%
    #     mutate(COUNTYNAME = convert_fips_to_names(COUNTYFP, states = rep("State", nrow(shp_join)), geo_header="COUNTY"))
    mutate(labels = paste0(
      "<strong> County: </strong> ",
      GEOID, "<br/> ",
      "<strong> Food Expenditure: </strong> $",
      fexpend_predictions, "<br/> ",
      "<strong> Food Insecurity: </strong> ",
      fsecurity_predictions, "<br/> "
    ) %>%
      lapply(htmltools::HTML))
})


#for color scale
limit <- c(min(final_combined$fexpend_predictions,na.rm=TRUE), max(final_combined$fexpend_predictions,na.rm=TRUE))
limit_bin <- c(min(final_combined$fsecurity_predictions,na.rm=TRUE), max(final_combined$fsecurity_predictions,na.rm=TRUE))
pal_bin <- colorBin(
  palette = "RdBu", domain = final_combined$fsecurity_predictions, 
  reverse = TRUE,
  bins = seq(0, max(final_combined$fsecurity_predictions, na.rm = TRUE) + .01, by = .20)
)
log_scale = seq(min(log(final_combined$fexpend_predictions), na.rm = TRUE), max(log(final_combined$fexpend_predictions), na.rm = TRUE), length.out = 11)
pal <- colorBin(
  palette = "RdBu", domain = final_combined$fexpend_predictions,
  reverse=FALSE,
  bins = exp(log_scale)
)

renderLeaflet({
  leaflet(shp_file2(),height=500, width=1000) %>%
    addTiles() %>%
    #setView(lng = 0, lat = 30, zoom = 2) %>%
    addPolygons(
      fillColor = ~ pal(fexpend_predictions),
      stroke=FALSE,
      color = "white",
      fillOpacity = 0.6,
      label = ~labels,
      highlight = highlightOptions(
        color = "black",
        bringToFront = TRUE
      )
    )%>%
    leaflet::addLegend(
      pal = pal, values = ~fexpend_predictions,
      opacity = 0.7, title = "Weekly Average\nFood Expenditure\nPer Person",
      position = "topright"
    )
})


```

Row
--------------------

### Predictions
```{r, echo = FALSE}
temp2 <- reactive({shp_file2() %>%as.data.frame%>% dplyr::select(c("GEOID","location.y", "fexpend_predictions","female.y", "hhsize.y", "hispanic.y","black.y", "kids.y", "elderly.y", "education.y", "employed.y","married.y", "disability.y", "urban_c.y"))%>%
                  rename(`Predicted Expenditure` = fexpend_predictions, `Location` = location.y, `Females per Household` = female.y,
                         `Hispanics per Household` = hispanic.y, `Blacks per Household` = black.y, `Kids per Household` = kids.y, 
                         `Elderly per Household` = elderly.y, `Educated per Household` = education.y, `Employed per Household` = employed.y,
                         `Married per Household` = married.y, `If Disabled in Household` = disability.y, `Urban level` = urban_c.y)
})

renderDataTable({temp2()})
```