---
title: "Comments on First Progress Report"
author: "Dr. Lendie Follett"
date: "6/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. disability plots: better labels on x and y axis (currently "disability" and "me")
2. disability plots: change so that the x axis is "yes" and "no" instead of numeric scale
3. What about a filled barchart like we did in 172? i.e., x axis = disability, fill by some measure of food insecurity (e.g., TRUE/FALSE if food insecurity > 0)
4. for the elderly plot on the left... should the x axis label be "number of elderly in household", not "food insecure"?
5. It's a zero inflated poisson so we have to say, in the Poisson interpretations, "For vulnerable individuals"...
6. In the Poisson interpretations, need to say "by a factor of" or use percent terminology. Basically, it's just like Poisson interpretations (see packet 3 from 172), but with the "For vulnerable individuals" intro.
7. We will interpret the logistic part of Zero-inflated Poisson just like we did in logistic regression (in terms of odds, see packet 2), but our "success" is being in the certain zero group. 
8. It's really really interesting that elderly has opposite effects in the bernoulli vs poisson components in the zero inflated model.
9. For your family size vs food expenses plot, try geom_jitter instead of geom_point to help reduce overplotting. Or, try boxplots with a binned version of hhsize. May have to create new variable for this.
10. labels on all plots
11. Try scale_y_log10() on some of the fexpend plots. That will log transform the y axis, while keeping the original scale on the labels. 
12. Might be work looking into the related-not related count data issue
13. With all disability plots - change from 0/1 to yes/no
14. Scale y axis on the fexpend plots to us dollar notation. Google "dollar axis labels ggplot2"
15. To make x axis clearer

```{r, eval=FALSE}
#change your_max to whatever the maximum mtry is
+scale_x_continuous(breaks = c(1:your_max), labels = c(1:your_max))
```

16. Color and overall beautifulness of plots. Color should always be useful, not just for pretties, but do try to include it. Recall how we got a list of color blind friendly palettes. 

